Kelp morphometric properties and the link to hydrodynamic modelling

### 2018-09-20 chapter_1.Rmd
Minor edits to chapter_1 script and time-frame added to proposal

The important environmental drivers of kelp idividuals and communities include light, substrata, salinity, sedimentation, nitrients, temperature and water motion. Although studies have investigated the effects of important environmental drivers, the roles these factors play is often difficult to evaluate as such factors may never be fully independent of each other, i.e. environmental factors are to some extent temperature dependent. Multifactorial studies have attempt to explain combined affects, however these studies are often limited to investigating combination of two or three environmental drivers as inclusion of too many factors can lead to results that are difficult to interept. Environmetal factors are highly variable on temporal and spatial scales, and their effects may also dependent on life-stage adding a further layer of complexity to investigations.

@@

#Depth

@@

#Nutrients

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment. Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity of macroalgae communties. The width, vertical zonation and diversity of algal communities often change predictably along gradients of exposure. This is also true for variation in morphological characters of individuals may also vary with wave exposure, such as frond width and degree of 'branching' [@moss1948]. @moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid in relation to wave exposure between study sites.

For example, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994]. @lewis found that degrees of wave exposure vertical structure rocky shore communities...**work in progress**...

>>>>

The important environmental drivers of kelp individuals and communities include light, substrata, salinity, sedimentation, nitrients, temperature and wave exposure. Although studies have investigated the effects of important environmental drivers, the roles these factors play is often difficult to evaluate as such factors may never be fully independent of each other, i.e. environmental factors are to some extent dependent on one another. Multifactorial studies have attempted to explain combined affects, however these studies are often limited to investigating combination of two or three environmental drivers as inclusion of too many factors can lead to results that are difficult to interept. Environmetal factors are highly variable on temporal and spatial scales and their effects may also be dependent on the life-stage of the organism, adding a further layer of complexity to investigations.

@@

#Substrata
#Salinity
#Depth

@@

#Sedimentation
#Nutrients

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment. Like temperature, wave exposure can act on regional scales and has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties. The width, vertical zonation and diversity of algal communities often change predictably along gradients of exposure. This is also true for variation in morphological characters of individuals may also vary with wave exposure, such as frond width and degree of 'branching' [@moss1948]. @moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid in relation to wave exposure between study sites.

Increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake through increased turbulence in the boundry layer [@lobban1994]. @lewis found that degrees of wave exposure vertical structure rocky shore communities...**work in progress**...

### 2018-09-20 proposal_draft_2018.Rmd
Minor edits to chapter_1 script and time-frame added to proposal

Advances in numerical modelling have been founded on physical wave theory on how a wave "behaves". This approach is based on a theoretical persepctive rather than the need to answer ecological questions. Besides diffraction, these models incorporate more complexity by including wind forcing, wave-to-wave interactions and loss of energy due to friction and wave breaking. Numerical models have a variety of applications and are often incorporated within hydrodynamic general circulation models and are used operationally for forecasting the sea state [@hasselmann1988; @booij1999; @smith2001]. The downside of advanced numerical models is that are computationally intensive which creates limitations for large scale simulations. Therefore, their application along long stretches of variable coastline, inshore environments and ocean-wide simulations limited due to the poor spatial coverage. However, this models can be designed for local or site specific coverage, provided the correct data is available.

@@

A combined quantitative ecology and oceanography approach will be used for this study. Morphological characteristics and their relation to the hydrodynamic environment must first be established and the findings will form the basis for the proceeding research. Once the relationship has been established it will be used to calibrate future hydrodynamic models developed during the course of the study. Once the hydrodynamic model has been established and calibrated it will be used to investigate the flow of microplastics along the South African coastline.

@@

>>>>

Advances in numerical modelling have been founded on physical wave theory on how a wave "behaves". This approach is based on a theoretical persepctive rather than the need to answer ecological questions. Besides diffraction, these models incorporate more complexity by including wind forcing, wave-to-wave interactions and loss of energy due to friction and wave breaking. Numerical models have a variety of applications and are often incorporated within hydrodynamic general circulation models and are used operationally for forecasting the sea state [@hasselmann1988; @booij1999; @smith2001]. The downside of advanced numerical models is that are computationally intensive which creates limitations for large scale simulations. Therefore, their application along long stretches of variable coastline, inshore environments and ocean-wide simulations are limited due to the poor spatial coverage. However, these models can be designed for local or site specific coverage, provided the correct data are available.

@@

A combined quantitative ecology and oceanography approach will be used for this study. Morphological characteristics and their relation to the hydrodynamic environment must first be established and the findings will form the basis for the proceeding chapters. Once the relationship has been established it will be used to calibrate future hydrodynamic models developed during the course of the study. Once the hydrodynamic model has been established and calibrated it will be used to investigate the flow of microplastics along the South African coastline.

@@

Months | Description
------------------------------ | -------------------------------------------
April 2018 - September 2018 | Develop and write proposal for submission to department
September 2018 - December 2018 | Acceptable level of progress on literature review. The first data chapter should also be complete and submission of first paper
January 2019 - May 2019 | Become competent in relevant software (Python). Data analyses for the 2nd data chapter should ideally be near completion. Substantial additions to literature review. 
May 2019 - August 2019| Completion of 2nd data chapter and relevant paper should also be near completion. 
August 2019 - December 2019 | Complete analyses and write-up for final data chapter. Significant progress on the relevant paper. Literature review should be near completion. 
January 2020 - April 2020 | This time will be used to make final edits and complete thesis write-up
